\assignment

### 2017-03-14 project/paper.rnw
Something like an intro.

\newcommand\assignment{Stat 534 Project: Point Process Intensity Surface Models}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{\today}

@@

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = FALSE, comment = NA, message = FALSE,
               fig.path = 'proj_figure/', cache.path = 'proj_cache/',
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 3,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

#library(xtable)
options(xtable.table.placement = 'H', width = 80, scipen = 2,
        show.signif.stars = FALSE)
@

@@

@@

\textcite{baddeleycoeurjolly14,baddeleyturner,diggleetal,bermanturner}

\section{Simple Example}

\section{Simulation}

>>>>

\newcommand\assignment{Stat 534 Project: \\
Extrapolation from Poisson Process Intensity Surface Models}
\newcommand\myname{Kenny Flagg}
\newcommand\duedate{\today}

@@

<<setup, echo = FALSE, message = FALSE, cache = FALSE>>=
library(knitr)
library(extrafont)
opts_chunk$set(echo = FALSE, comment = NA, message = FALSE,
               fig.path = 'paper_figure/', cache.path = 'paper_cache/',
               show.signif.stars = FALSE,
               fig.align = 'center', fig.width = 6.5, fig.height = 3,
               fig.pos = 'H', size = 'footnotesize', dev = 'pdf',
               dev.args = list(family = 'CM Roman', pointsize = 11))
knit_theme$set('print')

library(xtable)
options(xtable.table.placement = 'H', width = 80, scipen = 2,
        xtable.sanitize.rownames.function = function(x){return(x)},
        show.signif.stars = FALSE)

library(spatstat)

set.seed(83534)
@

@@

One little-studied use of inhomogeneous point process intensity estimation
is the use of an intensity surface model fit to events in a subregion to
predict the intensity over the entire region of interest. This procedure would
be relevant whenver it is known or suspected that some type of plant, animal,
or other item occurs in the region, and the goal of the analysis is to map the
trend in where these tend to be located rather than estimate parameters of
some process at work across hypothetical replicates of similar regions. It may
be prohibitively expensive or difficult to observe all events over the entire
region, so a sample of subregions is taken. For example, conservationists want
to study the spatial distribution of an endangered plant across a large region
of thick jungle so that they can establish a preserve where the plant is
protected. They cannot search the entire jungle, so they take a simple random
sample of quadrats and record the locations of the plants in the quadrats.
They will then want to fit a model describing the trends in intensity of these
plants extrapolated over the entire jungle. In this setting, the objective is
to map the realized spatial intensity of the plant over this jungle so that
these plants can be protected, not to estimate parameters of the process that
arranges plants of this species at jungles like this one.

Another situation where it is useful to extrapolate point pattern intensity
outside of the observed region is in mapping subsurface geomagnetic anomalies,
such as the munitions debris found at former military test ranges. This is
frequently done in the early stages of an unexploded ordnance (UXO)
remediation, where the intensity surface of inert munitions fragments is
used to identify the locations of targets so that the search for UXO can be
focused on the sections of the site most likely to contain it. The anomalies
are only observed when detection equipment passes directly over them, but the
project leaders are concerened with finding UXO and do not want to waste
resources finding every inert fragment that could be found by metal detectors.
The most common data collection method is to take a systematic sample of
 straight-line transects and observe anomalies in rectangular regions centered
along the transects. Frequently, moving averages of the intensity are computed
in circular windows and an intensity map is produced using ordinary kriging to
predict the intensity in a grid of these windows, but this assumes
stationarity when the moving averages are believed to be non-stationary,
ignores the point process nature of the data, and is very sensitive to the
window size~\parencite{vspguide,wp}. In this project, I explore the use of
polynomial trend surface models for the log-intensity at a simulated UXO site.

@@

\textcite{digglebook,bermanturner}

\section{Simple Examples}

log-linear example from HW4

simulate from
\begin{equation*}
\log(\lambda(x,y)) = 5x + 2y; \quad (x, y) \in (0, 1)^{2}
\end{equation*}
but do not observe any events in \((0.5, 0.8)^{2}\). Then fit
\begin{equation*}
\log(\lambda(x,y)) = \beta_{0} + \beta_{1}x + \beta_{2}y
\end{equation*}
and predict on \((-0.5, 1.5)^{2}\).

<<hw4_sim>>=
simulate_win <- owin(c(0, 1), c(0, 1))
observe_win <- setminus.owin(simulate_win, owin(c(0.5, 0.8), c(0.5, 0.8)))
predict_win <- owin(c(-0.5, 1.5), c(-0.5, 1.5))
hw4 <- rpoispp(function(x, y){exp(5 * x + 2 * y)}, win = simulate_win)
@

<<hw4_plot, fig.cap = 'One realization of a Poisson process with log-linear intensity on the unit square.'>>=
par(mar = c(0, 0, 1, 0))
plot(hw4, main = 'Event Locations')
@

<<hw4_fit_all>>=
hw4_fit_a <- ppm(hw4 ~ x + y, correction = 'isotropic')
hw4_pred_a <- predict(hw4_fit_a, window = predict_win, se = TRUE)
# By default predictions are on a 128 by 128 grid.
@
<<hw4_fit_unobserved>>=
hw4_fit_u <- ppm(hw4[observe_win] ~ x + y, correction = 'isotropic')
hw4_pred_u <- predict(hw4_fit_u, window = predict_win, se = TRUE)
@

<<hw4_fitcoefs_a, results = 'asis'>>=
hw4_coef_a <- summary(hw4_fit_a)$coefs.SE.CI[,c('Estimate', 'S.E.')]
rownames(hw4_coef_a) <- c('\\(\\widehat{\\beta_{0}}\\)',
                          '\\(\\widehat{\\beta_{1}}\\)',
                          '\\(\\widehat{\\beta_{2}}\\)')
xtable(hw4_coef_a, label = 'hw4esta',
       caption = 'Estimated coefficients for the log-linear trend model fit using the full region.')
@

<<hw4_fitcoefs_u, results = 'asis'>>=
hw4_coef_u <- summary(hw4_fit_u)$coefs.SE.CI[,c('Estimate', 'S.E.')]
rownames(hw4_coef_u) <- c('\\(\\widehat{\\beta_{0}}\\)',
                          '\\(\\widehat{\\beta_{1}}\\)',
                          '\\(\\widehat{\\beta_{2}}\\)')
xtable(hw4_coef_u, label = 'hw4estu',
       caption = 'Estimated coefficients for the log-linear trend model fit using a subset of the region.')
@

<<hw4_fitplot, fig.pos = 'p', fig.height = 8, fig.cap = 'Estimated log-intensity surface, log-scale prediction standard errors, and prediction coeffiecient of variation from models fit using all events in the full simulation region (left) and events in a subset of the simulation region (right). The regions and event locations are overlaid in white.'>>=
par(mfrow = c(3, 2), mar = c(0, 0, 1.2, 2), las = 2, cex = 1)

plot(log(hw4_pred_a$estimate), zlim = c(-4.5, 10.5),
     main = expression(log(hat(lambda)[f](italic(list(x,y))))))
plot(simulate_win, add = TRUE, border = '#ffffff40')
points(hw4, pch = '.', col = 'white')

plot(log(hw4_pred_u$estimate), zlim = c(-4.5, 10.5),
     main = expression(log(hat(lambda)[s](italic(list(x,y))))))
plot(observe_win, add = TRUE, border = '#ffffff40')
points(hw4[observe_win], pch = '.', col = 'white')

plot(log(hw4_pred_a$se), zlim = c(-4.5, 10.5),
     main = expression(log(SE(hat(lambda)[f](italic(list(x,y)))))))
plot(simulate_win, add = TRUE, border = '#ffffff40')
points(hw4, pch = '.', col = 'white')

plot(log(hw4_pred_u$se), zlim = c(-4.5, 10.5),
     main = expression(log(SE(hat(lambda)[s](italic(list(x,y)))))))
plot(observe_win, add = TRUE, border = '#ffffff40')
points(hw4[observe_win], pch = '.', col = 'white')

plot(hw4_pred_a$se/hw4_pred_a$estimate, zlim = c(0, 1),
     main = expression(SE(hat(lambda)[f](italic(list(x,y)))) /
                         hat(lambda)[f](italic(list(x,y)))))
plot(simulate_win, add = TRUE, border = '#ffffff40')
points(hw4, pch = '.', col = 'white')

plot(hw4_pred_u$se/hw4_pred_u$estimate, zlim = c(0, 1),
     main = expression(SE(hat(lambda)[s](italic(list(x,y)))) /
                         hat(lambda)[s](italic(list(x,y)))))
plot(observe_win, add = TRUE, border = '#ffffff40')
points(hw4[observe_win], pch = '.', col = 'white')
@

easy site, realization 2,000, 390 ft spacing

use \texttt{polynom()} not \texttt{poly()}!

\section{Simulation Study}

\section{Discussion and Conclusion}

smallish SEs outside observed region because of faith in the model form --
need model checking

how much of the region do you need to observe to trust your model?

\appendix
\section{R Code Appendix}
